\documentclass[11pt,a4paper]{article}

% ── Packages ──────────────────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Font: use default Computer Modern
\usepackage[english]{babel}
\usepackage[margin=2.2cm,top=2.4cm,bottom=2.4cm]{geometry}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{float}
\usepackage{multicol}
\usepackage{parskip}
\usepackage{caption}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, fit, backgrounds}

% ── Colors ────────────────────────────────────────────────────────────
\definecolor{rteblue}{HTML}{1B3A5C}
\definecolor{rtecyan}{HTML}{00A3E0}
\definecolor{rtegray}{HTML}{5A6872}
\definecolor{lightbg}{HTML}{F0F4F8}
\definecolor{accentgreen}{HTML}{2E8B57}

% ── Hyperref setup ────────────────────────────────────────────────────
\hypersetup{
  colorlinks=true,
  linkcolor=rteblue,
  citecolor=rteblue,
  urlcolor=rtecyan,
  pdfauthor={ST7 Cohort -- CentraleSupélec},
  pdftitle={LLM-Based Log Contextualization for Smart Grid Operations},
}

% ── Header/Footer ─────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{rtegray}{ST7 Smart Cities -- CentraleSupélec}}
\fancyhead[R]{\small\textcolor{rtegray}{March 2026}}
\fancyfoot[C]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ── Section styling ───────────────────────────────────────────────────
\titleformat{\section}{\large\bfseries\color{rteblue}}{\thesection}{0.8em}{}
\titleformat{\subsection}{\normalsize\bfseries\color{rteblue!80!black}}{\thesubsection}{0.6em}{}
\titlespacing*{\section}{0pt}{1.2em}{0.5em}
\titlespacing*{\subsection}{0pt}{0.8em}{0.3em}

% ── Compact lists ─────────────────────────────────────────────────────
\setlist{nosep, leftmargin=1.5em}

\begin{document}

% ══════════════════════════════════════════════════════════════════════
% TITLE PAGE
% ══════════════════════════════════════════════════════════════════════
\begin{titlepage}
\centering
\vspace*{1.5cm}

{\Huge\bfseries\color{rteblue} LLM-Based Log Contextualization\\[0.3em] for Smart Grid Operations\par}

\vspace{0.8cm}
{\Large\color{rtegray} Transforming Unstructured Operational Data\\into Actionable Insight\par}

\vspace{1.5cm}

\begin{tikzpicture}
  \fill[rteblue!10, rounded corners=8pt] (-4,-0.8) rectangle (4,0.8);
  \node[text=rteblue, font=\large\bfseries] at (0,0.15) {ST7 Smart Cities};
  \node[text=rtegray, font=\normalsize] at (0,-0.3) {CentraleSupélec -- March 2026};
\end{tikzpicture}

\vspace{1.5cm}

\begin{tabular}{rl}
  \textbf{Supervision:} & Noureddine Henka (RTE France) \\
  \textbf{Project Team:} & ST7 group \\
  \textbf{Academic Frame:} & Smart Cities Module \\
\end{tabular}

\vfill

{\small\color{rtegray} Project Report -- Asynchronous Multi-Agent Platform for Grid Log Analysis}
\end{titlepage}

% ══════════════════════════════════════════════════════════════════════
% BODY
% ══════════════════════════════════════════════════════════════════════

\section{Introduction and Context}

RTE (Réseau de Transport d'Électricité) operates the French high-voltage transmission grid, ensuring the continuous delivery of electricity from producers to consumers across the country. With the accelerating integration of intermittent renewable energy sources such as solar and wind, grid operations have become significantly more complex, requiring faster and more precise balancing decisions.

Grid operators rely on daily operational synthesis reports---dense, multi-page PDF documents written in French---to track incidents, outages, equipment faults, and market adjustments. These reports follow heterogeneous formatting conventions, may contain scanned pages, rasterized tables, and embedded diagrams, and demand substantial time to read and triage manually. During high-pressure operational periods, the latency introduced by manual reading is operationally unacceptable.

This project addresses the challenge by building an \textbf{asynchronous multi-agent platform} that automatically ingests these PDF logs, extracts structured incident data, classifies and prioritizes incidents by severity, and produces a decision-ready analysis report enriched with domain knowledge via Retrieval-Augmented Generation (RAG). The platform automates extraction and triage into actionable, RAG-grounded markdown output, with end-to-end latency depending on document quality, OCR load, and LLM response time.

The project was carried out within the ST7 Smart Cities module at CentraleSupélec under the supervision of RTE France, with three formal deliverables: a state-of-the-art review on LLMs and RAG for operational contextualization, a fully implemented pipeline with prompt engineering and RTE log integration, and OCR-assisted extraction support for scanned and rasterized operational PDF pages.

\section{State of the Art: LLMs and RAG for Grid Operations}

Large Language Models such as GPT-4 and Mistral have demonstrated strong capabilities in information extraction and zero-shot reasoning, acting as general-purpose logic engines. However, applying them directly to domain-specific operational contexts like electricity grid management presents a fundamental limitation: pure LLMs lack native knowledge of internal grid codes, RTE-specific procedures, and operational terminology, which makes them prone to hallucination when generating domain-specific analysis.

Retrieval-Augmented Generation (RAG) addresses this gap by injecting retrieved domain knowledge directly into the LLM prompt at inference time. In this implementation, the approach follows a four-step cycle: incident context is converted to a retrieval query over a lightweight JSON vector index seeded from historical PDF documents; relevant chunks are retrieved based on semantic similarity; these chunks are merged into an enriched prompt; and the LLM generates a grounded response. This architecture improves domain grounding without requiring expensive model fine-tuning, making it suitable for operational deployment where procedures evolve frequently.

The literature review conducted as part of this project confirmed that the combination of deterministic preprocessing (rule-based extraction) with selective LLM invocation offers the best trade-off between reliability, auditability, and coverage of edge cases in document formats.

\section{System Architecture}

The platform follows a clear separation between a \textbf{frontend} user interface and a \textbf{backend} processing engine, connected through a RESTful API with asynchronous job management.

\subsection{Frontend (Next.js Dashboard)}

The frontend is a Next.js (React) application providing a PDF upload interface, periodic job status polling via \texttt{GET /api/jobs/\{job\_id\}}, stage-by-stage pipeline progress visualization (collector, preprocessing, incident, analysis), markdown report rendering, and an LLM diagnostics panel exposing provider, model, latency, parse status, and raw output preview. Users can export the generated analysis as Markdown, plain text, or print-to-PDF directly from the interface.

\subsection{Backend (FastAPI Engine)}

The backend is a FastAPI service that validates uploads, stores PDF files durably, creates and tracks asynchronous jobs in persistent SQL storage, executes the multi-agent orchestration pipeline in background tasks, and persists intermediate stage artifacts for traceability. The API returns both the final analysis markdown and detailed stage history for frontend rendering.

\subsection{Orchestrator and Job Processor}

The orchestrator is the central coordinator for strict sequential execution of the four pipeline agents: \texttt{collector $\rightarrow$ preprocessing $\rightarrow$ incident $\rightarrow$ analysis}. It instantiates all agents against a shared repository, emits stage updates (running, completed, failed), captures payloads and errors for each stage, and returns \texttt{partial} when preprocessing or incident fails after earlier success, while analysis failure returns \texttt{failed} with upstream stage payloads preserved.

The job processor wraps the orchestrator with runtime reliability concerns: background execution, concurrency control via semaphore, timeout and retry policy, and persistent job lifecycle updates (\texttt{queued}, \texttt{running}, \texttt{completed}, \texttt{partial}, \texttt{failed}). In practice: \textit{Orchestrator = what steps run and in what order}, \textit{JobProcessor = how jobs run safely in production}.

\subsection{Data and Knowledge Layer}

Persistent storage is SQL-backed (SQLite for local development and PostgreSQL for deployment) for job metadata, stage snapshots, and artifacts. Uploaded files are stored on the filesystem with SHA-256 content hashing that enables deduplication-aware job reuse. RAG retrieval is implemented through backend retriever components backed by a lightweight JSON vector index seeded from historical PDF documents.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  node distance=0.6cm and 1.2cm,
  box/.style={rectangle, draw=rteblue, fill=rteblue!8, rounded corners=4pt,
              minimum height=0.85cm, minimum width=2.8cm, align=center,
              font=\small\bfseries, text=rteblue},
  db/.style={cylinder, draw=rtegray, fill=lightbg, shape border rotate=90,
             aspect=0.25, minimum height=0.9cm, minimum width=1.6cm,
             font=\scriptsize, text=rtegray, align=center},
  arr/.style={-{Stealth[length=5pt]}, thick, rteblue!70},
  darr/.style={-{Stealth[length=4pt]}, dashed, rtegray},
]
  % Frontend
  \node[box, fill=rtecyan!12, draw=rtecyan] (fe) {Next.js\\Dashboard};
  % Backend agents
  \node[box, right=1.8cm of fe] (col) {Collector};
  \node[box, below=of col] (pre) {Preprocessing};
  \node[box, below=of pre] (inc) {Incident};
  \node[box, below=of inc] (ana) {Analysis};
  % Databases
  \node[db, right=1.6cm of col, yshift=-0.8cm] (pdb) {SQL Store\\(SQLite/PostgreSQL)};
  \node[db, right=1.6cm of inc] (vdb) {JSON Vector Index\\(RAG)};

  % Arrows
  \draw[arr] (fe) -- node[above, font=\scriptsize, text=rtegray]{REST API} (col);
  \draw[arr] (col) -- (pre);
  \draw[arr] (pre) -- (inc);
  \draw[arr] (inc) -- (ana);
  \draw[darr] (col.east) -- ++(0.6,0) |- (pdb.west);
  \draw[darr] (ana.east) -- ++(0.4,0) |- (vdb.west);
  % Backend box
  \begin{scope}[on background layer]
    \node[draw=rteblue!40, fill=rteblue!3, rounded corners=6pt,
          fit=(col)(pre)(inc)(ana), inner sep=10pt,
          label={[font=\scriptsize\bfseries, text=rteblue]above:FastAPI \& Custom Orchestrator}] {};
  \end{scope}
\end{tikzpicture}
\caption{High-level architecture: Frontend, Backend Pipeline, and Data Layer.}
\label{fig:arch}
\end{figure}

\section{End-to-End Data Flow}

\begin{enumerate}
  \item \textbf{Upload}: The frontend uploads a PDF via \texttt{POST /api/analyze}.
  \item \textbf{Job Creation}: The backend stores the file, creates a job ID, and starts asynchronous processing.
  \item \textbf{Collector}: Raw text is extracted page-by-page, with OCR fallback quality checks.
  \item \textbf{Preprocessing}: Report metadata, sections, and normalized events are built.
  \item \textbf{Incident}: Events are enriched with severity, tags, escalation, and priority ordering.
  \item \textbf{Analysis}: Deterministic statistics plus LLM/RAG synthesis produce the final markdown report.
  \item \textbf{Rendering}: The frontend polls \texttt{GET /api/jobs/\{job\_id\}} and displays timeline, report, and diagnostics.
\end{enumerate}

\section{Agent Pipeline: Detailed Design}

\subsection{Agent 1 -- Collector (Ingestion \& Extraction)}

The Collector agent ingests the uploaded PDF and produces per-page raw text output with extraction quality signals. It computes a SHA-256 hash for content identity and deduplication support, extracts text page-by-page using native PDF extraction (\texttt{pypdf}), flags weak pages (short text / low word count), and attempts OCR fallback (\texttt{pypdfium2} + \texttt{pytesseract}) when needed. For each page, it stores the selected text, extraction method (\texttt{pypdf}, \texttt{ocr}, or \texttt{pypdf+ocr}), and whether downstream fallback is still required.

\subsection{Agent 2 -- Preprocessing (Structuring Data)}

The Preprocessing agent transforms raw page text into structured report metadata, document sections, and candidate events. Deterministic parsing extracts report title/date/scope metadata (language is currently fixed to \texttt{fr} in the schema), highlighted bullet sections, section boundaries with page ranges, event type, time range, impact values (MW/customers), location and asset hints, cause categories, and operational actions. LLM fallback is used only for weak pages flagged by Collector, with a strict JSON prompt and normalization into the shared schema while preserving French wording when possible. Rule-based and LLM-extracted events are merged and deduplicated before persistence.

\subsection{Agent 3 -- Incident Management (Triage)}

The Incident agent converts extracted events into actionable incidents with severity, tags, escalation status, and a sorted priority queue. Severity combines customer and MW thresholds with contextual rules (critical infrastructure mentions and high-risk keywords such as intrusion or supervision loss). It then derives operational tags (for example \texttt{customer\_outage}, \texttt{industrial\_impact}, \texttt{telecom\_loss}, \texttt{third\_party\_damage}, \texttt{malicious\_act}, \texttt{media\_sensitive}) and marks incidents for escalation when severity is \texttt{critical} or \texttt{high}.

\subsection{Agent 4 -- Analysis (LLM + RAG Synthesis)}

The Analysis agent produces document-level intelligence and canonical markdown output. First, it computes deterministic statistics (incident count, severity/cause/region/asset distributions, total MW and customer impact), recurring patterns, repeated assets, and caveats. It then retrieves contextual chunks via the RAG retriever and calls the LLM with a strict JSON schema. Post-processing applies quality gates, normalizes list fields, and merges deterministic fallback insights/actions with LLM output. When LLM generation and parsing succeed, the markdown report is rendered into fixed sections: Incident Analysis, Executive Summary, Key Statistics, Patterns, Cross-Incident Insights, Recommended Actions, Reasoning Summary, and Caveats; otherwise the analysis stage is marked failed and diagnostics are returned.

\section{Technical Stack}

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Layer} & \textbf{Technologies} \\
\midrule
Frontend & Next.js 16 (React 19), Tailwind CSS 4, shadcn/ui, react-markdown, react-dropzone \\
Backend \& Orchestration & FastAPI (Python 3.12), SQLAlchemy (async), custom orchestrator with stage callbacks \\
AI / ML Engine & Configurable LLM service (Hugging Face Router or local Ollama), strict JSON prompting, diagnostics capture, hashing-based embeddings \\
PDF/OCR Extraction & pypdf, pypdfium2, pytesseract (French + English OCR) \\
Data \& Storage & SQL persistence (SQLite dev / PostgreSQL deployment), JSON-based vector store, durable file storage, SHA-256 deduplication \\
DevOps & Docker Compose, GitHub Actions CI (lint, build, test), Kubernetes manifests \\
\bottomrule
\end{tabularx}
\caption{Technical stack matrix across all system layers.}
\label{tab:stack}
\end{table}

\section{Results and Demonstration}

The implemented system demonstrates the complete asynchronous workflow from PDF upload to final report rendering: job creation, stage-level execution tracking, structured event extraction, severity-based incident triage, and markdown analysis generation. The frontend successfully visualizes per-stage status and exposes LLM diagnostics (provider/model/latency/parse status/error preview), enabling transparent operator review.

Backend test coverage includes pipeline orchestration behavior, incident scoring, preprocessing fallback logic, analysis-stage normalization/quality gates, API job contracts, and an opt-in smoke execution on real PDF logs (\texttt{RUN\_REAL\_PDF\_TESTS=1}). This validates that the four operational stages (\texttt{collector}, \texttt{preprocessing}, \texttt{incident}, \texttt{analysis}) are implemented end-to-end in code.

\section{Limitations and Future Work}

The current system has three principal constraints. First, OCR performance degrades on low-quality scanned pages and dense industrial log layouts where text is embedded in complex graphics or rasterized tables. Second, the RAG index requires curated historical documents, creating a maintenance burden. Third, the system relies on general-purpose foundation models (via Hugging Face or Ollama) without dedicated fine-tuning on RTE-specific vocabulary and procedures.

Future development directions include transitioning from batch PDF uploads to real-time log streaming for continuous monitoring, implementing automated alert routing directly into existing RTE SCADA/EMS systems, and deploying a smaller domain-fine-tuned model trained specifically on RTE's historical operational corpus to improve extraction accuracy and reduce inference costs.

\subsection{Presentation Accuracy Note}

The implemented backend pipeline currently contains four execution agents: \texttt{collector}, \texttt{preprocessing}, \texttt{incident}, and \texttt{analysis}. In conceptual diagrams, an additional ``Indexer Management Agent'' may appear; in the current codebase, indexing/retrieval is handled inside the RAG components used by \texttt{AnalysisAgent}, not as a standalone orchestrator stage.

\section{Conclusion}

This project delivers an asynchronous multi-agent platform that transforms raw grid-operation PDF logs into structured incidents, risk-prioritized triage, and decision-ready analysis reports for operators. By combining deterministic preprocessing for speed and auditability with selective LLM invocation for edge-case coverage, and grounding generated insights through RAG retrieval of historical domain documents, the system supports operator decision workflows with traceable outputs. Operator triage effort is reduced through automation, with latency depending on input quality and model/infrastructure configuration.

\end{document}
